Summary of tumor cell classification with `AUCell` for `r params$library_id`

### 2024-07-12 analyses/cell-type-ewings/template_notebooks/auc-workflow/01-auc-results.Rmd
add plotting of marker gene cutoff

```{r}
# generate classification df to use for plots 
classification_df <- sce |> 
  scuttle::makePerCellDF(use.dimred = "UMAP") |> 
  # replace UMAP.1 with UMAP1
  dplyr::rename_with(
    \(x) stringr::str_replace(x, "^UMAP\\.", "UMAP")
  ) |>
  # get rid of excess columns
  dplyr::select(barcodes, UMAP1, UMAP2, singler_celltype_annotation) |> 
  # join with previous annotations, singler results, and gene set scores 
  dplyr::left_join(auc_results_df, by = "barcodes") |> 
  dplyr::left_join(geneset_scores_df, by = "barcodes") |> 
  dplyr::mutate(
    # make a factor so tumor always appears first 
    auc_classification = forcats::fct_relevel(auc_classification, "Tumor")
  )

@@

@@

```{r}
ggplot(classification_df, aes(x = UMAP1, y = UMAP2, color = auc_classification)) +
  geom_point(size = 0.5, alpha = 0.5) +
  scale_color_manual(values = plot_colors) +
  labs(
    color = "AUCell classification"
  )
```

@@

```{r, eval=has_marker_gene_exp}
# create a density plot showing the distribution of marker gene expression across classification methods 
markers_df |> 
  dplyr::select(barcodes, sum_raw_exp, auc_classification) |> 
  dplyr::distinct() |> 
  ggplot(aes(x = sum_raw_exp, color = auc_classification)) +
  geom_density() +
  scale_color_manual(values = plot_colors) +
  labs(
    x = "Raw sum of marker gene logcounts",
    color = "AUCell classification"
  )
```

@@

annotation <- ComplexHeatmap::columnAnnotation(
  AUCell = classification_df$auc_classification, 
  col = list(AUCell = plot_colors)
)
```

@@

```{r}
# prep data frame for plotting gene set scores 
geneset_plot_df <- classification_df |> 
  dplyr::select(barcodes, auc_classification, starts_with("mean-")) |> 
  # make one column containing all gene set scores 
  tidyr::pivot_longer(
    cols = starts_with("mean"),
    names_to = "geneset",
    values_to = "mean_score"
  ) |> 
  # simplify gene set name for labels
  dplyr::mutate(
    geneset = stringr::word(geneset, -1, sep = "-")
  )

# make density plot showing distribution of gene set scores for tumor/normal 
ggplot(geneset_plot_df, aes(x = mean_score, color = auc_classification)) +
  geom_density(bw = 0.05) +
  facet_grid(rows = vars(geneset),
             scales = "free_y") +
  scale_color_manual(values = plot_colors) +
  labs(
    x = "Gene set score",
    color = "AUCell classification"
  )
```

>>>>

```{r}
# generate classification df to use for plots 
classification_df <- sce |> 
  scuttle::makePerCellDF(use.dimred = "UMAP") |> 
  # replace UMAP.1 with UMAP1
  dplyr::rename_with(
    \(x) stringr::str_replace(x, "^UMAP\\.", "UMAP")
  ) |>
  # get rid of excess columns
  dplyr::select(barcodes, UMAP1, UMAP2, singler_celltype_annotation) |> 
  # join with previous annotations, singler results, and gene set scores 
  dplyr::left_join(auc_results_df, by = "barcodes") |> 
  dplyr::left_join(geneset_scores_df, by = "barcodes")

@@

```{r}
# add marker gene classification classifying cells based on expression of at least 1 marker gene  
marker_gene_class_df <- markers_df |> 
  dplyr::select(barcodes, sum_raw_exp) |> 
  dplyr::distinct() |> 
  dplyr::mutate(
    marker_gene_classification = dplyr::if_else(
      sum_raw_exp > 0,
      "Tumor",
      "Normal"

    )
  )

classification_df <- classification_df |> 
  dplyr::left_join(marker_gene_class_df, by = "barcodes") |> 
  tidyr::pivot_longer(ends_with("classification"),
                      names_to = "method",
                      values_to = "classification") |> 
  dplyr::mutate(
    # make a factor so tumor always appears first 
    classification = forcats::fct_relevel(classification, "Tumor")
  )
```

@@

```{r}
ggplot(classification_df, aes(x = UMAP1, y = UMAP2, color = classification)) +
  geom_point(size = 0.5, alpha = 0.5) +
  facet_wrap(vars(method)) +
  scale_color_manual(values = plot_colors) +
  labs(
    color = ""
  ) +
  guides(color = guide_legend(override.aes = list(size = 3, alpha = 1)))
```

@@

```{r, eval=has_marker_gene_exp}
# create a density plot showing the distribution of marker gene expression across classification methods 
classification_df |> 
  ggplot(aes(x = sum_raw_exp, color = classification)) +
  geom_density() +
  facet_wrap(vars(method),
             scales = "free_y") +
  scale_color_manual(values = plot_colors) +
  labs(
    x = "Raw sum of marker gene logcounts",
    color = ""
  ) 
```

@@

annotation_df <- classification_df |> 
  dplyr::select(barcodes, classification, method) |> 
  tidyr::pivot_wider(names_from = method, 
                     values_from = classification)

annotation <- ComplexHeatmap::columnAnnotation(
  AUCell = annotation_df$auc_classification, 
  marker_genes = annotation_df$marker_gene_classification,
  col = list(AUCell = plot_colors,
             marker_genes = plot_colors)
)
```

@@

```{r}
# prep data frame for plotting gene set scores 
geneset_plot_df <- classification_df |> 
  dplyr::select(barcodes, classification, method, starts_with("mean-")) |> 
  # make one column containing all gene set scores 
  tidyr::pivot_longer(
    cols = starts_with("mean"),
    names_to = "geneset",
    values_to = "mean_score"
  ) |> 
  # simplify gene set name for labels
  dplyr::mutate(
    geneset = stringr::word(geneset, -1, sep = "-")
  )

# make density plot showing distribution of gene set scores for tumor/normal 
ggplot(geneset_plot_df, aes(x = mean_score, color = classification)) +
  geom_density(bw = 0.05) +
  facet_grid(rows = vars(geneset),
             cols = vars(method),
             scales = "free_y") +
  scale_color_manual(values = plot_colors) +
  labs(
    x = "Gene set score",
    color = ""
  )
```
